---
title: "North Sea SMS Keyrun Supplemental Plots"
author: "Sarah Gaichas, Valerio Bartolino, Xochitl Cormon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   html_document:
     code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggthemes)
library(RColorBrewer)

```

```{r spplist}
# load the species list (checked that they are the same for the 2020 and 2017 runs)
spp <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2023/species_names.in", nrows=27)
spp <- gsub("_","",spp$V1)
spp <- data.frame(Species.n=1:length(spp), stkName=spp)

```

## Observed Data vs SMS Predicted

### Catch {.tabset}

#### 2023 SMS KeyRun
```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Obs vs Pred catch (here only 2023 run)
dat <- read.table("https://github.com/ices-eg/wg_WGSAM/raw/master/NorthSeaKeyRun_2023/summary_table_raw.out", sep="", header=T , na.strings="", stringsAsFactors=F)
#dat[1:3,]

dat <- left_join(dat,spp) %>%
     mutate(Yield = as.numeric(as.character(SOP)),
            Yield.hat = as.numeric(as.character(SOP.hat)),
            Year = as.numeric(as.character(Year))) %>%
     gather("Source","Yield",9:10)

#postscript(paste(dirFigs,"catch_obsVSpre.ps",sep="/"))
ggplot(dat) +
    geom_point(aes(Year,Yield,col=Source)) +
    geom_line(aes(Year,Yield,col=Source,lty=Source)) +
    facet_wrap(~stkName, scales="free_y")+
    xlim(1973,2023)+
    theme_tufte() +
    theme(legend.position="bottom")
#dev.off()

```


#### 2020 SMS KeyRun
```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Obs vs Pred catch (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary_table_raw.out", sep="", header=T , na.strings="", stringsAsFactors=F)
#dat[1:3,]

dat <- left_join(dat,spp) %>%
    mutate(Yield = as.numeric(as.character(SOP)),
           Yield.hat = as.numeric(as.character(SOP.hat)),
           Year = as.numeric(as.character(Year))) %>%
    gather("Source","Yield",8:9)

#postscript(paste(dirFigs,"catch_obsVSpre.ps",sep="/"))
ggplot(dat) +
    geom_point(aes(Year,Yield,col=Source)) +
    geom_line(aes(Year,Yield,col=Source,lty=Source)) +
    facet_wrap(~stkName, scales="free_y")+
    xlim(1973,2023)+
    theme_tufte() +
    theme(legend.position="bottom")
#dev.off()

```

#### 2017 SMS KeyRun
```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Obs vs Pred catch (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2017/summary_table_raw.out", sep="", header=T , na.strings="", stringsAsFactors=F)
#dat[1:3,]

dat <- left_join(dat,spp) %>%
    mutate(Yield = as.numeric(as.character(Yield)),
           Yield.hat = as.numeric(as.character(Yield.hat)),
           Year = as.numeric(as.character(Year))) %>%
    gather("Source","Yield",8:9)

#postscript(paste(dirFigs,"catch_obsVSpre.ps",sep="/"))
ggplot(dat) +
    geom_point(aes(Year,Yield,col=Source)) +
    geom_line(aes(Year,Yield,col=Source,lty=Source)) +
    facet_wrap(~stkName, scales="free_y")+
    xlim(1973,2023)+
    theme_tufte() +
    theme(legend.position="bottom")
#dev.off()

```

### {-}

### Diet proportion 2023 

```{r, message=FALSE, warning=FALSE}
# plot Obs vs Pred stomachs (here only 2020 run)
dat <- read.table("https://github.com/ices-eg/wg_WGSAM/raw/master/NorthSeaKeyRun_2023/summary_stom.out", header=T)
#dat[1:4,]

sppPred <- spp %>% mutate(Predator.no = Species.n) %>% rename(PredName=stkName)
sppPrey <- spp %>% mutate(Prey.no = Species.n) %>% rename(PreyName=stkName)

dat <- dat %>%
    left_join(sppPred) %>%
    mutate(Species.n = NULL) %>%
    left_join(sppPrey) %>%
    mutate(PreyName = as.character(PreyName),
           PreyName = ifelse(is.na(PreyName), "otherfood", PreyName))

tmp <- dat %>%
    group_by(Year,PredName,PreyName) %>%
    summarise(stomcon = sum(stomcon,na.rm=T),
              stomcon.hat = sum(stomcon.hat,na.rm=T))

# pie plot Obs and Pred
preys <- as.character(unique(tmp$PreyName))
preds <- as.character(unique(tmp$PredName))
#plList <- vector("list",length(preds))
nb.cols <- 14
colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

plist2 = lapply(split(tmp, tmp$PredName), function(d) {
#for(i in 1:length(preds)){
    ggplot(d %>% gather("source","stomcon",4:5)) +
    geom_bar(aes(x="",stomcon,fill=PreyName), width = 1, stat = "identity", position="fill") +
    scale_fill_manual(values=colPalette) +
    facet_grid(source~Year) +
    #ggtitle(preds[[i]]) +
    coord_polar("y")
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v1.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

# same thing as barplot plot Obs and Pred
#for(i in 1:length(preds)){
plist3 = lapply(split(tmp, tmp$PredName), function(d) {
  ggplot(d) +
    geom_col(aes(Year-0.2,stomcon,fill=PreyName), width=0.35, stat="identity", position="stack") +
    geom_col(aes(Year+0.2,stomcon.hat,fill=PreyName), width=0.35, stat="identity", position="stack") +
    scale_fill_manual(values=colPalette) +
    #ggtitle(preds[[i]]) +
    xlab("Year") +
    theme_bw()
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v2.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

```

### Pie charts {.tabset}

```{r pieloop23, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist2[preds[i]]) 
  cat("  \n")
}
```

### {-}

### Bar charts {.tabset}

```{r barloop23, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist3[preds[i]]) 
  cat("  \n")
}
```


### {-}


### Diet proportion 2020 

```{r, message=FALSE, warning=FALSE}
# plot Obs vs Pred stomachs (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary_stom.out", header=T)
#dat[1:4,]

sppPred <- spp %>% mutate(Predator.no = Species.n) %>% rename(PredName=stkName)
sppPrey <- spp %>% mutate(Prey.no = Species.n) %>% rename(PreyName=stkName)

dat <- dat %>%
    left_join(sppPred) %>%
    mutate(Species.n = NULL) %>%
    left_join(sppPrey) %>%
    mutate(PreyName = as.character(PreyName),
           PreyName = ifelse(is.na(PreyName), "otherfood", PreyName))

tmp <- dat %>%
    group_by(Year,PredName,PreyName) %>%
    summarise(stomcon = sum(stomcon,na.rm=T),
              stomcon.hat = sum(stomcon.hat,na.rm=T))

# pie plot Obs and Pred
preys <- as.character(unique(tmp$PreyName))
preds <- as.character(unique(tmp$PredName))
#plList <- vector("list",length(preds))
nb.cols <- 14
colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

plist2 = lapply(split(tmp, tmp$PredName), function(d) {
#for(i in 1:length(preds)){
    ggplot(d %>% gather("source","stomcon",4:5)) +
    geom_bar(aes(x="",stomcon,fill=PreyName), width = 1, stat = "identity", position="fill") +
    scale_fill_manual(values=colPalette) +
    facet_grid(source~Year) +
    #ggtitle(preds[[i]]) +
    coord_polar("y")
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v1.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

# same thing as barplot plot Obs and Pred
#for(i in 1:length(preds)){
plist3 = lapply(split(tmp, tmp$PredName), function(d) {
  ggplot(d) +
    geom_col(aes(Year-0.2,stomcon,fill=PreyName), width=0.35, stat="identity", position="stack") +
    geom_col(aes(Year+0.2,stomcon.hat,fill=PreyName), width=0.35, stat="identity", position="stack") +
    scale_fill_manual(values=colPalette) +
    #ggtitle(preds[[i]]) +
    xlab("Year") +
    theme_bw()
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v2.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

```

### Pie charts {.tabset}

```{r pieloop, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist2[preds[i]]) 
  cat("  \n")
}
```

### {-}

### Bar charts {.tabset}

```{r barloop, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist3[preds[i]]) 
  cat("  \n")
}
```


### {-}

## SMS output comparisons

### Numbers at age in Q1 {.tabset}

```{r Nagecomp, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Num@age from 2017 and 2020 runs
dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2017/summary.out", sep="", header=T , na.strings="", stringsAsFactors=F)
dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary.out", sep="", header=T , na.strings="", stringsAsFactors=F)
dat23 <- read.table("https://github.com/ices-eg/wg_WGSAM/raw/master/NorthSeaKeyRun_2023/summary.out", sep="", header=T , na.strings="", stringsAsFactors=F)


dat17 <- left_join(dat17,spp) %>% mutate(run="r17")
dat20 <- left_join(dat20,spp) %>% mutate(run="r20")
dat23 <- left_join(dat23,spp) %>% mutate(run="r23")
dat <- bind_rows(dat17,dat20, dat23)

# select Herring in Q1
#i <- spp$stkName[21] # 21 is herring
#postscript(paste(dirFigs,"/numAtAge_",i,".ps",sep=""))
plist = lapply(split(dat, dat$stkName), function(d) {
  ggplot(d %>% filter(Quarter == 1)) +
    geom_point(aes(Year,N, col=run)) +
    geom_line(aes(Year,N, col=run)) +
    facet_wrap(~Age, scale="free_y")+
    theme_tufte() +
    theme(legend.position="bottom")
})
#dev.off()

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist$Herring
```

#### Mackerel
```{r,  fig.width=9, fig.height=8}
plist$Mackerel
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=6}
plist$'N.sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=3}
plist$'Nor.pout'
```

#### Saithe
```{r,  fig.width=9, fig.height=8}
plist$'Saithe'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=6}
plist$'S.sandeel'
```

#### Sprat
```{r,  fig.width=9, fig.height=3}
plist$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist$Whiting
```

### {-}

```{r dietcolor}
# plot part M2 per predator
dat23 <- read.table("https://github.com/ices-eg/wg_WGSAM/raw/master/NorthSeaKeyRun_2023/Input_Output/Output/WhoEatsWhom/who_eats_whom_level1.csv", sep=",", header=T)

dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/who_eats_whom_level1.csv", sep=",", header=T)
#dat20 <- read.csv("C://Users//xocor//Documents//Work//Projects//WGSAM//who_eats_whom_level1.csv", sep=",", header=T, dec=".",stringsAsFactors=FALSE)
#head(dat20)

dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/Input_Output/Output/WhoEatsWhom/who_eats_whom_level1.csv", sep=",", header=T)
dat17 <- dat17 %>% mutate(run="r17")
dat20 <- dat20 %>% mutate(run="r20")
dat23 <- dat23 %>% mutate(run="r23")
dat2 <- bind_rows(dat17,dat20, dat23)

# aggregate birds
dat3 <- dat2 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","GBB.Gull","Guillemot","Her. Gull","Her.Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator),
           Predator = ifelse(Predator %in% c("A. radiata", "A.radiata"), "A.radiata", Predator),
           Predator = ifelse(Predator %in% c("G. gurnards", "G.gurnards"), "G.gurnards", Predator),
           Predator = ifelse(Predator %in% c("Grey seal" , "Grey.seal"), "Grey.seal", Predator),
           Predator = ifelse(Predator %in% c("H. porpoise", "H.porpoise"), "H.porpoise", Predator),
           Predator = ifelse(Predator %in% c("N.horse mac", "N.horse.mac"), "N.horse.mac", Predator),
           Predator = ifelse(Predator %in% c("W.horse mac", "W.horse.mac"), "W.horse.mac", Predator),
           Prey = ifelse(Prey %in% c("N. sandeel", "N.sandeel"), "N.sandeel", Prey),
           Prey = ifelse(Prey %in% c("Nor. pout",  "Nor.pout"), "Nor.pout", Prey),
           Prey = ifelse(Prey %in% c("S. sandeel", "S.sandeel"), "S.sandeel", Prey)
           ) %>%
    group_by(Year,Quarter,Predator, Predator.age,Prey, Prey.age, run) %>%
    summarise(M2 = sum(Part.M2, na.rm=T)) %>%
	ungroup()
	
# length(unique(dat3$Predator)) #http://medialab.github.io/iwanthue/
col <- c("#000047","#858a00","#ff2b47","#00d3c9", "#0188d2", "#7426d6","#e37b00","#ffa0ee","#930025","#00bd3b","yellow","black","#005144")
names(col) <- as.factor(unique(dat3$Predator))

preycol <- c("#da62e7", "#549700", "#00609a", "#8dd971", "#535622", "lightgray")
names(preycol) <- as.factor(c("N.sandeel","Nor.pout","S.sandeel","Herring","Sprat","Plaice"))

col <- c(col, preycol)

```

### Modeled Q1 predator diet {.tabset}

#### 2023 SMS KeyRun

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# plot predators model diet (here only 2020 run)
dat23 <- read.table("https://github.com/ices-eg/wg_WGSAM/raw/master/NorthSeaKeyRun_2023/Input_Output/Output/WhoEatsWhom/who_eats_whom_level2.csv", sep=",", header=T)
#dat[1:4,]

# aggregate birds
datQ23 <- dat23 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Guillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
    group_by(Year,Quarter,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

# All predators diet in Q1
#postscript(paste(dirFigs,"diet_byPredator_Q1.ps",sep="/"))
ggplot(datQ23 %>% filter(Quarter == 1)) +
    geom_bar(aes(Year,eatenW,fill=Prey), stat="identity", position="fill") +
    facet_wrap(~Predator)+
    scale_fill_manual(values=col) +
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()

```


#### 2020 SMS KeyRun

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# plot predators model diet (here only 2020 run)
dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/who_eats_whom_level2.csv", sep=",", header=T)
#dat[1:4,]

# aggregate birds
datQ20 <- dat20 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Guillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
    group_by(Year,Quarter,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

# All predators diet in Q1
#postscript(paste(dirFigs,"diet_byPredator_Q1.ps",sep="/"))
ggplot(datQ20 %>% filter(Quarter == 1)) +
    geom_bar(aes(Year,eatenW,fill=Prey), stat="identity", position="fill") +
    facet_wrap(~Predator)+
    scale_fill_manual(values=col) +
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()

```

#### 2017 SMS KeyRun

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# plot predators model diet **check to make sure this is 2017 output**
dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/Input_Output/Output/WhoEatsWhom/who_eats_whom_level2.csv", sep=",", header=T)
#dat[1:4,]

# aggregate birds
datQ17 <- dat17 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Guillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
    group_by(Year,Quarter,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

# All predators diet in Q1
#postscript(paste(dirFigs,"diet_byPredator_Q1.ps",sep="/"))
ggplot(datQ17 %>% filter(Quarter == 1)) +
    geom_bar(aes(Year,eatenW,fill=Prey), stat="identity", position="fill") +
    facet_wrap(~Predator)+
    scale_fill_manual(values=col) +
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()

```

### {-}

### Modeled predators contributing to predation by prey {.tabset}

#### 2023 SMS proportion by weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# ------------------------
# plot predators contributing to predation by prey
preys <- as.character(unique(datQ23$Prey))
plList <- vector("list",length(preys))
#nb.cols <- 14
#colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)


# # by quarter
# for(i in 1:length(plList)){
# plList[[i]] <- ggplot(datQ %>% filter(Prey == preys[i])) +
#     geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
#     scale_fill_manual(values=colPalette) +
#     facet_wrap(~Quarter) +
#     ggtitle(preys[i])
# }
# for(i in 1:length(plList)){
# postscript(paste(dirFigs,"/predators_on_",preys[i],".ps",sep=""))
# print(plList[[i]])
# dev.off()
# }

# by year
datY <- datQ23 %>%
    group_by(Year,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

#postscript(paste(dirFigs,"predators_on_eachprey.ps",sep="/"))
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()
```

#### 2023 SMS weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    #ylim(0,2e+06)+
    theme(legend.position="bottom")


```

#### 2020 SMS proportion by weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# ------------------------
# plot predators contributing to predation by prey
preys <- as.character(unique(datQ20$Prey))
plList <- vector("list",length(preys))
#nb.cols <- 14
#colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)


# # by quarter
# for(i in 1:length(plList)){
# plList[[i]] <- ggplot(datQ %>% filter(Prey == preys[i])) +
#     geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
#     scale_fill_manual(values=colPalette) +
#     facet_wrap(~Quarter) +
#     ggtitle(preys[i])
# }
# for(i in 1:length(plList)){
# postscript(paste(dirFigs,"/predators_on_",preys[i],".ps",sep=""))
# print(plList[[i]])
# dev.off()
# }

# by year
datY <- datQ20 %>%
    group_by(Year,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

#postscript(paste(dirFigs,"predators_on_eachprey.ps",sep="/"))
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()
```

#### 2020 SMS weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    #ylim(0,2e+06)+
    theme(legend.position="bottom")


```

#### 2017 SMS proportion by weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# ------------------------
# plot predators contributing to predation by prey
preys <- as.character(unique(datQ17$Prey))
plList <- vector("list",length(preys))
#nb.cols <- 14
#colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

# # by quarter
# for(i in 1:length(plList)){
# plList[[i]] <- ggplot(datQ %>% filter(Prey == preys[i])) +
#     geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
#     scale_fill_manual(values=colPalette) +
#     facet_wrap(~Quarter) +
#     ggtitle(preys[i])
# }
# for(i in 1:length(plList)){
# postscript(paste(dirFigs,"/predators_on_",preys[i],".ps",sep=""))
# print(plList[[i]])
# dev.off()
# }

# by year
datY <- datQ17 %>%
    group_by(Year,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

#postscript(paste(dirFigs,"predators_on_eachprey.ps",sep="/"))
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    theme(legend.position="bottom")
#dev.off()
```

#### 2017 SMS weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity") +
    #scale_fill_manual(values=colPalette) +
    scale_fill_manual(values=col) +
    facet_wrap(~Prey)+
    xlim(1973,2023)+
    #ylim(0,2e+06)+
    theme(legend.position="bottom")


```


### {-}

### Modeled M2 by predator on prey {.tabset}

```{r, fig.width=9, fig.height=8, message=FALSE, results="asis"}
# plot part M2 per predator
#dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/who_eats_whom_level1.csv", sep=",", header=T)
#dat20 <- read.csv("C://Users//xocor//Documents//Work//Projects//WGSAM//who_eats_whom_level1.csv", sep=",", header=T, dec=".",stringsAsFactors=FALSE)
#head(dat20)

#dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/Input_Output/Output/WhoEatsWhom/who_eats_whom_level1.csv", sep=",", header=T)
#dat17 <- dat17 %>% mutate(run="r17")
#dat20 <- dat20 %>% mutate(run="r20")
#dat2 <- bind_rows(dat17,dat20)

# aggregate birds
#dat3 <- dat2 %>%
#    mutate(Predator = as.character(Predator),
#           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Guillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
#    group_by(Year,Quarter,Predator, Predator.age,Prey, Prey.age, run) %>%
#    summarise(M2 = sum(Part.M2, na.rm=T)) %>%
#	ungroup()
	
# length(unique(dat3$Predator)) #http://medialab.github.io/iwanthue/
#col <- c("#000047","#858a00","#ff2b47","#00d3c9", "#0188d2", "#7426d6","#e37b00","#ffa0ee","#930025","#00bd3b","yellow","black","#005144")
#names(col) <- as.factor(unique(dat3$Predator))

dat3 <- subset(dat3, Year <=2017)

#pdf("C://Users//xocor//Documents//Work//Projects//WGSAM//M2.pdf", pointsize=10)
for (s in c("Herring", "Whiting", "Sprat")){
  for(x in c(1:max(dat3[dat3$Prey==s,"Prey.age"]))){
    g <- ggplot(dat3 %>% subset(Prey==s&Prey.age==x)) +
      geom_bar(aes(Year,M2,fill=Predator), stat="identity", position="fill") +
      facet_grid(Quarter~run)+ scale_fill_manual(values=col) + ggtitle(paste(s, "Age", x, sep=" "))
    #windows()
    cat("  \n####",  s, " Age ",x,"  \n")
    print(g) 
    cat("  \n")
  }}
#dev.off()


```

### {-}