---
title: "North Sea Herring M2 Comparson"
author: "Sarah Gaichas"
date: "10/13/2020"
output: 
   html_document:
     code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggthemes)

```

## 2017 vs 2020 Herring M2 from North Sea SMS

Output figures comparing M2 by predator (should match Morten's outputs)

```{r readdat, message=FALSE}
M2_2017 <- read_csv(here("KeyRunComparisons" , "NorthSeaSMS2017", "who_eats_whom_level1.csv"))
M2_2020 <- read_csv(here("KeyRunComparisons" , "NorthSeaSMS2020", "who_eats_whom_level1.csv"))
M2_2020_old <- read_csv(here("KeyRunComparisons" , "NorthSeaSMS2020", "who_eats_whom_level1_old.csv"))
```

### 2017 Keyrun M2 (fixed y axis) {.tabset}
```{r plotM2017}

plist = lapply(split(M2_2017, M2_2017$Prey), function(d) {
  ggplot(d, aes(Year, Part.M2, fill=Predator)) + 
    geom_bar(stat = "identity") +
    facet_wrap(Prey~Prey.age) +
    xlab("Year") +
    ylab("M2") +
    theme_tufte() +
    theme(legend.position="bottom")
})

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist$Herring
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=8}
plist$'N. sandeel'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=8}
plist$'S. sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=8}
plist$'Nor. pout'
```

#### Sprat
```{r,  fig.width=9, fig.height=8}
plist$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist$Whiting
```

### {-}

### 2020 Keyrun M2 (fixed y axis) {.tabset}
```{r plotM2020}

plist2 = lapply(split(M2_2020, M2_2020$Prey), function(d) {
  ggplot(d, aes(Year, Part.M2, fill=Predator)) + 
    geom_bar(stat = "identity") +
    facet_wrap(Prey~Prey.age) +
    xlab("Year") +
    ylab("M2") +
    theme_tufte() +
    theme(legend.position="bottom")
})

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist2$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist2$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist2$Herring
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=8}
plist2$'N. sandeel'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=8}
plist2$'S. sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=8}
plist2$'Nor. pout'
```

#### Sprat
```{r,  fig.width=9, fig.height=8}
plist2$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist2$Whiting
```

### 2017 Keyrun M2 (variable y axis) {.tabset}
```{r plotM20172}

plist3 = lapply(split(M2_2017, M2_2017$Prey), function(d) {
  ggplot(d, aes(Year, Part.M2, fill=Predator)) + 
    geom_bar(stat = "identity") +
    facet_wrap(Prey~Prey.age, scales = "free_y") +
    xlab("Year") +
    ylab("M2") +
    theme_tufte() +
    theme(legend.position="bottom")
})

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist3$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist3$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist3$Herring
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=8}
plist3$'N. sandeel'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=8}
plist3$'S. sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=8}
plist3$'Nor. pout'
```

#### Sprat
```{r,  fig.width=9, fig.height=8}
plist3$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist3$Whiting
```

### {-}

### 2020 Keyrun M2 (variable y axis) {.tabset}
```{r plotM20202}

plist4 = lapply(split(M2_2020, M2_2020$Prey), function(d) {
  ggplot(d, aes(Year, Part.M2, fill=Predator)) + 
    geom_bar(stat = "identity") +
    facet_wrap(Prey~Prey.age, scales = "free_y") +
    xlab("Year") +
    ylab("M2") +
    theme_tufte() +
    theme(legend.position="bottom")
})

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist4$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist4$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist4$Herring
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=8}
plist4$'N. sandeel'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=8}
plist4$'S. sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=8}
plist4$'Nor. pout'
```

#### Sprat
```{r,  fig.width=9, fig.height=8}
plist4$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist4$Whiting
```

### {-}

## Compare Herring M2 between 2017 and 2020 

### Now tabs are herring predators {.tabset}

```{r compareM, message=FALSE, warning=FALSE}

M2_2017 <- add_column(M2_2017, KeyRun = "SMS2017") 
M2_2020 <- add_column(M2_2020, KeyRun = "SMS2020")
M2_2020_old <- add_column(M2_2020_old, KeyRun = "SMS2020_old")
M2_comp <- bind_rows(M2_2017, M2_2020)
M2_comp <- bind_rows(M2_comp, M2_2020_old)

M2_compHerring <- M2_comp %>%
  filter(Prey=="Herring") %>%
  group_by(Year, Predator, Prey, Prey.age, KeyRun) %>%
  summarise_at(vars(eatenW, Part.M2), funs(sum))

plist5 = lapply(split(M2_compHerring, M2_compHerring$Predator), function(d) {
  ggplot(d, aes(Year, Part.M2, group=KeyRun)) + 
    geom_point(aes(colour=KeyRun)) +
    geom_line()+
    facet_wrap(Prey~Prey.age, scales = "free_y", ncol = 2) +
    xlab("Year") +
    ylab("M2") +
    theme_tufte() +
    theme(legend.position="bottom")
})
```

#### Cod
```{r,  fig.width=9, fig.height=9}
plist5$Cod
```

#### Fulmar
```{r,  fig.width=9, fig.height=3}
plist5$Fulmar
```

#### G. gurnards
```{r,  fig.width=9, fig.height=3}
plist5$'G. gurnards'
```

#### Gannet
```{r,  fig.width=9, fig.height=9}
plist5$Gannet
```

#### GBB. Gull
```{r,  fig.width=9, fig.height=6}
plist5$'GBB. Gull'
```

#### Grey seal
```{r,  fig.width=9, fig.height=9}
plist5$'Grey seal'
```

#### Guillemot
```{r,  fig.width=9, fig.height=3}
plist5$Guillemot
```

#### H. porpoise
```{r,  fig.width=9, fig.height=9}
plist5$'H. porpoise'
```

#### Hake
```{r,  fig.width=9, fig.height=8}
plist5$Hake
```

#### Her. Gull
```{r,  fig.width=9, fig.height=6}
plist5$'Her. Gull'
```

#### Kittiwake
```{r,  fig.width=9, fig.height=6}
plist5$Kittiwake
```

#### Mackerel
```{r,  fig.width=9, fig.height=3}
plist5$Mackerel
```

#### N.horse mac
```{r,  fig.width=9, fig.height=3}
plist5$'N.horse mac'
```

#### Puffin
```{r,  fig.width=9, fig.height=3}
plist5$Puffin
```

#### Razorbill
```{r,  fig.width=9, fig.height=3}
plist5$Razorbill
```

#### Saithe
```{r,  fig.width=9, fig.height=9}
plist5$Saithe
```

#### Whiting
```{r,  fig.width=9, fig.height=9}
plist5$Whiting
```

### {-}

Find biggest M2 differences for age 2 herring between key-runs:

```{r herrM2diff}

M2_diffHerring <- M2_compHerring %>%
  arrange(Year, Predator, Prey, Prey.age, KeyRun) %>% 
  group_by(Year, Predator, Prey, Prey.age) %>% 
  mutate(diffM2_2020 = c(diff(Part.M2), NA),
         diffeatenW_2020 = c(diff(eatenW), NA)) %>%
  filter(!is.na(diffM2_2020))

age2diff <- M2_diffHerring %>%
  filter(Prey.age==2) %>%
  arrange(diffM2_2020)

library(DT)
datatable(age2diff, rownames = FALSE, options = list(pageLength = 25))


```

