---
title: "North Sea SMS Keyrun Supplemental Plots"
author: "Valerio Bartolino, Sarah Gaichas"
date: "10/14/2020"
output: 
   html_document:
     code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggthemes)
library(RColorBrewer)

```

```{r spplist}
# load the species list (checked that they are the same for the 2020 and 2017 runs)
spp <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2017/species_names.in", nrows=27)
spp <- gsub("_","",spp$V1)
spp <- data.frame(Species.n=1:length(spp), stkName=spp)

```

## Observed Data vs SMS Predicted

### Catch {.tabset}

#### 2020 SMS KeyRun
```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Obs vs Pred catch (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary_table_raw.out", sep="", header=T , na.strings="", stringsAsFactors=F)
#dat[1:3,]

dat <- left_join(dat,spp) %>%
    mutate(Yield = as.numeric(as.character(Yield)),
           Yield.hat = as.numeric(as.character(Yield.hat)),
           Year = as.numeric(as.character(Year))) %>%
    gather("Source","Yield",8:9)

#postscript(paste(dirFigs,"catch_obsVSpre.ps",sep="/"))
ggplot(dat) +
    geom_point(aes(Year,Yield,col=Source)) +
    geom_line(aes(Year,Yield,col=Source,lty=Source)) +
    facet_wrap(~stkName, scales="free_y")+
    xlim(1973,2020)+
    theme_tufte() +
    theme(legend.position="bottom")
#dev.off()

```

#### 2017 SMS KeyRun
```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Obs vs Pred catch (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2017/summary_table_raw.out", sep="", header=T , na.strings="", stringsAsFactors=F)
#dat[1:3,]

dat <- left_join(dat,spp) %>%
    mutate(Yield = as.numeric(as.character(Yield)),
           Yield.hat = as.numeric(as.character(Yield.hat)),
           Year = as.numeric(as.character(Year))) %>%
    gather("Source","Yield",8:9)

#postscript(paste(dirFigs,"catch_obsVSpre.ps",sep="/"))
ggplot(dat) +
    geom_point(aes(Year,Yield,col=Source)) +
    geom_line(aes(Year,Yield,col=Source,lty=Source)) +
    facet_wrap(~stkName, scales="free_y")+
    xlim(1973,2020)+
    theme_tufte() +
    theme(legend.position="bottom")
#dev.off()

```

### {-}

### Diet proportion 2020 

```{r, message=FALSE, warning=FALSE}
# plot Obs vs Pred stomachs (here only 2020 run)
dat <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary_stom.out", header=T)
#dat[1:4,]

sppPred <- spp %>% mutate(Predator.no = Species.n) %>% rename(PredName=stkName)
sppPrey <- spp %>% mutate(Prey.no = Species.n) %>% rename(PreyName=stkName)

dat <- dat %>%
    left_join(sppPred) %>%
    mutate(Species.n = NULL) %>%
    left_join(sppPrey) %>%
    mutate(PreyName = as.character(PreyName),
           PreyName = ifelse(is.na(PreyName), "otherfood", PreyName))

tmp <- dat %>%
    group_by(Year,PredName,PreyName) %>%
    summarise(stomcon = sum(stomcon,na.rm=T),
              stomcon.hat = sum(stomcon.hat,na.rm=T))

# pie plot Obs and Pred
preys <- as.character(unique(tmp$PreyName))
preds <- as.character(unique(tmp$PredName))
#plList <- vector("list",length(preds))
nb.cols <- 14
colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

plist2 = lapply(split(tmp, tmp$PredName), function(d) {
#for(i in 1:length(preds)){
    ggplot(d %>% gather("source","stomcon",4:5)) +
    geom_bar(aes(x="",stomcon,fill=PreyName), width = 1, stat = "identity", position="fill") +
    scale_fill_manual(values=colPalette) +
    facet_grid(source~Year) +
    #ggtitle(preds[[i]]) +
    coord_polar("y")
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v1.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

# same thing as barplot plot Obs and Pred
#for(i in 1:length(preds)){
plist3 = lapply(split(tmp, tmp$PredName), function(d) {
  ggplot(d) +
    geom_col(aes(Year-0.2,stomcon,fill=PreyName), width=0.35, stat="identity", position="stack") +
    geom_col(aes(Year+0.2,stomcon.hat,fill=PreyName), width=0.35, stat="identity", position="stack") +
    scale_fill_manual(values=colPalette) +
    #ggtitle(preds[[i]]) +
    xlab("Year") +
    theme_bw()
})
# for(i in 1:length(plList)){
#     postscript(paste(dirFigs,"/stomachs_fitting_",preds[i],"_v2.ps",sep=""))
#     print(plList[[i]])
#     dev.off()
# }

```

### Pie charts {.tabset}

```{r pieloop, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist2[preds[i]]) 
  cat("  \n")
}
```

### {-}

### Bar charts {.tabset}

```{r barloop, fig.width=9, message=FALSE, results="asis"}

for(i in 1:length(preds)) {
  cat("  \n####",  preds[i],"  \n")
  print(plist3[preds[i]]) 
  cat("  \n")
}
```


### {-}

## SMS output comparisons

### Numbers at age in Q1 {.tabset}

```{r Nagecomp, message=FALSE, warning=FALSE, fig.width=9, fig.height=8}
# plot Num@age from 2017 and 2020 runs
dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/NorthSeaKeyRun_2017/summary.out", sep="", header=T , na.strings="", stringsAsFactors=F)
dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/summary.out", sep="", header=T , na.strings="", stringsAsFactors=F)

dat17 <- left_join(dat17,spp) %>% mutate(run="r17")
dat20 <- left_join(dat20,spp) %>% mutate(run="r20")
dat <- bind_rows(dat17,dat20)

# select Herring in Q1
#i <- spp$stkName[21] # 21 is herring
#postscript(paste(dirFigs,"/numAtAge_",i,".ps",sep=""))
plist = lapply(split(dat, dat$stkName), function(d) {
  ggplot(d %>% filter(Quarter == 1)) +
    geom_point(aes(Year,N, col=run)) +
    geom_line(aes(Year,N, col=run)) +
    facet_wrap(~Age, scale="free_y")+
    theme_tufte() +
    theme(legend.position="bottom")
})
#dev.off()

```

#### Cod
```{r,  fig.width=9, fig.height=8}
plist$Cod
```

#### Haddock
```{r,  fig.width=9, fig.height=8}
plist$Haddock
```

#### Herring
```{r,  fig.width=9, fig.height=8}
plist$Herring
```

#### Mackerel
```{r,  fig.width=9, fig.height=8}
plist$Mackerel
```

#### N. Sandeel
```{r,  fig.width=9, fig.height=6}
plist$'N.sandeel'
```

#### Norway pout
```{r,  fig.width=9, fig.height=3}
plist$'Nor.pout'
```

#### Saithe
```{r,  fig.width=9, fig.height=8}
plist$'Saithe'
```

#### S. Sandeel
```{r,  fig.width=9, fig.height=6}
plist$'S.sandeel'
```

#### Sprat
```{r,  fig.width=9, fig.height=3}
plist$Sprat
```

#### Whiting
```{r,  fig.width=9, fig.height=8}
plist$Whiting
```

### {-}

### Modeled Q1 predator diet {.tabset}

#### 2020 SMS KeyRun

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# plot predators model diet (here only 2020 run)
dat20 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/KeyRunComparisons/NorthSeaSMS2020/who_eats_whom_level2.csv", sep=",", header=T)
#dat[1:4,]

# aggregate birds
datQ20 <- dat20 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Gillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
    group_by(Year,Quarter,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

# All predators diet in Q1
#postscript(paste(dirFigs,"diet_byPredator_Q1.ps",sep="/"))
ggplot(datQ20 %>% filter(Quarter == 1)) +
    geom_bar(aes(Year,eatenW,fill=Prey), stat="identity", position="fill") +
    facet_wrap(~Predator)+
    xlim(1973,2020)+
    theme(legend.position="bottom")
#dev.off()

```

#### 2017 SMS KeyRun

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# plot predators model diet **check to make sure this is 2017 output**
dat17 <- read.table("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/Input_Output/Output/WhoEatsWhom/who_eats_whom_level2.csv", sep=",", header=T)
#dat[1:4,]

# aggregate birds
datQ17 <- dat17 %>%
    mutate(Predator = as.character(Predator),
           Predator = ifelse(Predator %in% c("Fulmar","Gannet","GBB. Gull","Gillemot","Her. Gull","Kittiwake","Puffin","Razorbill"), "Birds", Predator)) %>%
    group_by(Year,Quarter,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

# All predators diet in Q1
#postscript(paste(dirFigs,"diet_byPredator_Q1.ps",sep="/"))
ggplot(datQ17 %>% filter(Quarter == 1)) +
    geom_bar(aes(Year,eatenW,fill=Prey), stat="identity", position="fill") +
    facet_wrap(~Predator)+
    xlim(1973,2020)+
    theme(legend.position="bottom")
#dev.off()

```

### {-}

### Modeled predators contributing to predation by prey {.tabset}

#### 2020 SMS proportion

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# ------------------------
# plot predators contributing to predation by prey
preys <- as.character(unique(datQ20$Prey))
plList <- vector("list",length(preys))
nb.cols <- 14
colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

# # by quarter
# for(i in 1:length(plList)){
# plList[[i]] <- ggplot(datQ %>% filter(Prey == preys[i])) +
#     geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
#     scale_fill_manual(values=colPalette) +
#     facet_wrap(~Quarter) +
#     ggtitle(preys[i])
# }
# for(i in 1:length(plList)){
# postscript(paste(dirFigs,"/predators_on_",preys[i],".ps",sep=""))
# print(plList[[i]])
# dev.off()
# }

# by year
datY <- datQ20 %>%
    group_by(Year,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

#postscript(paste(dirFigs,"predators_on_eachprey.ps",sep="/"))
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
    scale_fill_manual(values=colPalette) +
    facet_wrap(~Prey)+
    xlim(1973,2020)+
    theme(legend.position="bottom")
#dev.off()
```

#### 2020 SMS weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity") +
    scale_fill_manual(values=colPalette) +
    facet_wrap(~Prey)+
    xlim(1973,2020)+
    #ylim(0,2e+06)+
    theme(legend.position="bottom")


```

#### 2017 SMS proportion

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
# ------------------------
# plot predators contributing to predation by prey
preys <- as.character(unique(datQ17$Prey))
plList <- vector("list",length(preys))
nb.cols <- 14
colPalette <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)

# # by quarter
# for(i in 1:length(plList)){
# plList[[i]] <- ggplot(datQ %>% filter(Prey == preys[i])) +
#     geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
#     scale_fill_manual(values=colPalette) +
#     facet_wrap(~Quarter) +
#     ggtitle(preys[i])
# }
# for(i in 1:length(plList)){
# postscript(paste(dirFigs,"/predators_on_",preys[i],".ps",sep=""))
# print(plList[[i]])
# dev.off()
# }

# by year
datY <- datQ17 %>%
    group_by(Year,Predator,Prey) %>%
    summarise(eatenW = sum(eatenW, na.rm=T))

#postscript(paste(dirFigs,"predators_on_eachprey.ps",sep="/"))
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity", position="fill") +
    scale_fill_manual(values=colPalette) +
    facet_wrap(~Prey)+
    xlim(1973,2020)+
    theme(legend.position="bottom")
#dev.off()
```

#### 2017 SMS weight

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=8}
ggplot(datY) +
    geom_bar(aes(Year,eatenW,fill=Predator), stat="identity") +
    scale_fill_manual(values=colPalette) +
    facet_wrap(~Prey)+
    xlim(1973,2020)+
    #ylim(0,2e+06)+
    theme(legend.position="bottom")


```


### {-}
