---
title: "Seal diet analysis"
author: "Vanessa Trijoulet"
output: bookdown::word_document2
#always_allow_html: true
bibliography: biblio.bib
---


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(readxl)

sps<-  matrix(c(
  "COD",	"Gadus morhua",	'cod',"Cod",
  "WHG",	"Merlangius merlangus", 'whiting',"Whiting",
  "HAD",	"Melanogrammus aeglefinus", "haddock","Haddock",
  "POK",	"Pollachius virens","saithe","Saithe",
  "MAC",	"Scomber scombrus","mackerel","Mackerel",
  "HER",	"Clupea harengus","herring","Herring",
  "SAN",	"Ammodytidae","sandeel",'N.sandeel',
  "NOP",	"Trisopterus esmarkii","norway pout","Nor.pout",
  "SPR",	"Sprattus sprattus","sprat",'Sprat',
  "PLE",  "Pleuronectes platessa","plaice","Plaice",
  "SOL",	"Solea solea","sole","Sole"),
  byrow=T,ncol=4)
colnames(sps)<-c('SMS','latin','species',"Species")
sps<-as.data.frame(sps)

## Diet files
datadir <-"data/"
files <- list.files(datadir)
files <- files[grep("Hg cons ", files)]

cons.q.area <- list()
for (fil in files){
## Extract total consumption per quarter
  ss<-read_xlsx(path=file.path(datadir, fil),sheet="cons") %>% rename(prey=species.name)
  ss
  cons.q.area[[fil]] <- by(ss,ss$quarter,function(x){
    a<-xtabs(cons.t~prey+region,data=x)
    round(cbind(a,rowSums(a)),2)
  }) # in tonnes
nameCons <- lapply(cons.q.area[[fil]], rownames)[[1]]
  # Change naming so same as SMS
  nameCons[which(nameCons %in% sps$Species)] <- sort(sps[which(sps$Species %in% nameCons),]$SMS)
  nameCons[which(nameCons=="Norway pout")] <- "NOP"
  nameCons[which(nameCons=="Sandeel")] <- "SAN"
  for (k in 1:length(cons.q.area[[fil]])) rownames(cons.q.area[[fil]][[k]])=nameCons
}
```


# Data available

Grey seal diet was made available for the years 1985, 2002, and 2010-2011 by the SMRU (Phil Hammond pers. comm.).
This includes seal consumption per fish stock included in SMS (`r paste(sps[which(sps$SMS %in% nameCons),"species"], collapse=", ")`) in tonnes per year, quarter (Q1-4), and regions of the North Sea (regions 1-4, Shetland, Orkney and northern North Sea, central North Sea, and southern North Sea respectively). It also includes  outputs from otolith experiments as estimated fish length of fish in the diet given otolith size (per year, region, and quarter). The analysis of seal diet is described in detail in @hammond2016. These data allow to allocate total consumption to length class of fish prey.

# Method to allocate length distribution to grey seal consumption estimates

Each fish length sample is converted to weight using length-weight relationship parameters from @coull1989 (used also in the previous keyruns). The fish weights are summed across regions so that the weight consumed is given per species, length bins, quarters, and year. Weights are then converted to proportion consumed per length bin, and these proportions are multiplied by the total grey seal consumption per species and quarter to obtained the weight of prey consumed per length bin. The biomass of other food eaten by grey seals is derived from the total grey seal consumption per quarter and year.
The output csv file ("adjusted_seal_diet.csv") contains the biomass of prey eaten by all grey seals (column "prey_w") per length bin, quarter, and year.
The number of scat samples per quarter and year is used to give information on uncertainty in the diet data, those are added to the column "n_food" in the output diet file.

Two assumptions were made while handling grey seal diet, as follows:  
- Sprat was added to other food because of the small total consumption in each year and the lack of length samples (was already other food in the 2020 keyrun).  
- If there is less than 5 length bins for a prey in one quarter and year, the length distribution from the adjacent quarter is added to these samples. The borrowing is made between quarters 1-2 and quarters 3-4.

The code  for the analysis is available on the [WGSAM GitHub repository](https://github.com/ices-eg/wg_WGSAM/blob/master/NorthSeaKeyRun_2023/Seal_diet/seal_2023.R). The script creates figures of the length distribution before and after borrowing.


# Method to split sandeel diet per area (northern or southern North Sea)

Sandeel in the North Sea area managed as six individually assessed stocks. Given the lack of input data at the correct spatial scale, two sandeel stocks are considered in SMS and split into northern and southern North Sea stocks. In the previous SMS keyruns, the total grey seal predation was attributed to the northern sandeel stock. In the 2023 keyrun, proportion of sandeel consumed by grey seals was extracted from the diet data with the assumption that the northern areas correspond to Shetland, Orkney and northern North Sea, and the southern area to the central North Sea, and the southern North Sea. This resulted in the proportions in Table \@ref(tab:propSandeel). These proportions are used to split the diet data between northern and southern sandeel.

```{r propSandeel, echo=FALSE, warning=FALSE, message=FALSE}
# From consumption of sandeel
# Assumption North = Shetland, Orkney and northern North Sea, South = central North Sea, and southern North Sea
tmp <- lapply(cons.q.area, function(y) sapply(y, function(x) x["SAN", c("1","2","3","4")])) # row=region, col=quarter
tmp2 <- lapply(tmp, function(x) cbind("North" = apply(x[c("1","2"),],2,sum)/apply(x,2,sum), "South" = apply(x[c("3","4"),],2,sum)/apply(x,2,sum)))
propSandeel <-cbind(tmp2[[1]], tmp2[[2]], tmp2[[3]])
colnames(propSandeel) <- paste(c(rep("1985 ",2), rep("2002 ",2), rep("2010 ",2)), colnames(propSandeel))
rownames(propSandeel) <- paste("Quarter ", 1:4)

knitr::kable(propSandeel, digits=3, caption = "Proportion of sandeel consumed per area")
#x %>% add_header_above(c('1985'=2,'2002'=2,'2010'=2))
#x %>% as_image()
```



# References